Kernel Square-Loss Exemplar Machines

### 2016-11-15 sec_kernel_methods.tex
version J-1

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution of (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

Note that Eq. (\ref{beta:first}) allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq. (\ref{eq:first}) by setting $y_i=-1$ for all but one training sample. For our approach, we wish to solve (\ref{beta:first}) for many positive training samples (one at a time) against the same set of negative training samples. In the following subsections, we show how to take advantage of the fixed negative samples to efficiently solve (\ref{beta:first}).

\subsection{Offline preprocess of negative samples}\label{offline}
In order to calculate offline all operations that are dependent only on negative samples, let us suppose that the positive training sample is unknown and $K$ is the kernel matrix of the negative samples $X$.
The preprocessing phase consists of the calculation of the decomposition $B$ and the constants of Eq. (\ref{Wood:omega}): $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
These operations are done offline and their results are stored. In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
%Equation (\ref{beta:first}) has the same structure of Eq. (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.

@@

This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Eq. (\ref{beta:first}) now reads
\begin{align}
 \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. Solution $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
%This can be expressed by the linear system

\subsection{Similarity score}\label{simi_score}
Once the optimal parameters $(\beta, \nu)$ from (\ref{beta:final}) and the coordinates $u$, $v$ of $b_0'$ from (\ref{eq:lemma1}) have been found\footnote{We drop the ``$\star$'' in this subsection to avoid cluttering the notation.}, they can be used directly for measuring similarity between matching images.

Indeed, suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
We write $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. 
Moreover, $\alpha$ can be expressed by the linear system
\begin{equation}
\begin{bmatrix} \alpha_0 \\ \hat{\alpha} \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0 \\ \hat{\beta} \end{bmatrix}.\label{alpha:to:beta}
\end{equation}
Using Eq. (\ref{alpha:to:beta}) and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\lambda^{-2}(k(x_0,x_0')-v^T).
		%& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^T).
\end{split}
\end{equation}

For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM. Each image depends therefore on a vector of $p+2r$ dimensions \PP{Remind what are $p$ and $r$, I lost the former!}.

>>>>

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution of Eq. (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

Note that Eq. (\ref{beta:first}) is written as the usual form of a linear classifier. In particular, it allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq. (\ref{eq:first}) by setting $y_i=-1$ for all but one training sample. For our approach, we wish to solve (\ref{beta:first}) for many positive training samples (one at a time) against the same set of negative training samples. In the following subsections, we show how to take advantage of the fixed negative samples to efficiently solve (\ref{beta:first}).

\subsection{Offline preprocessing of negative samples}\label{offline}
Let us now return to the (kernelized) SLEM with $l$ as the square-loss.
In order to calculate offline all operations that are dependent only on negative samples, let us suppose that the positive training sample is unknown and $K$ is the kernel matrix of the negative samples $X$.
The preprocessing phase consists of the calculation of the decomposition $B$ and the constants of Eq. (\ref{Wood:omega}): $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
These operations are done offline and their results are stored. 
%In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
%Equation (\ref{beta:first}) has the same structure of Eq. (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.

@@

This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Eq. (\ref{beta:first}) now reads
\begin{align}
 \dfrac{1}{n}(b_0'^T\beta+\nu - 1)^2 +\dfrac{1}{n}\sumi (b_i'^T\beta+\nu + 1)^2
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Eq. (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. The solution $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
%This can be expressed by the linear system

\subsection{Similarity score}\label{simi_score}
%It is important to note that the exemplar classifier $h$ is never explicitly calculated, but we are able to compute the similarity between a pair of exemplar classifiers regardless.
Once the optimal parameters $(\beta, \nu)$ from (\ref{beta:final}) and the coordinates $u$, $v$ of $b_0'$ from (\ref{eq:lemma1}) have been found\footnote{We drop the ``$\star$'' in this subsection to avoid cluttering the notation.}, they can be used directly for measuring similarity between matching images.

Suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
We write $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. 
Moreover, $\alpha$ can be expressed by the linear system
\begin{equation}
\begin{bmatrix} \alpha_0 \\ \hat{\alpha} \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0 \\ \hat{\beta} \end{bmatrix}.\label{alpha:to:beta}
\end{equation}
Using Eq. (\ref{alpha:to:beta}) and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\lambda^{-2}(k(x_0,x_0')-v^T).
		%& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^T).
\end{split}
\end{equation}

For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its similarity score to whichever other image for SLEM. Since we assume the base feature $x_0$ has dimension $p$ and $\hat{\beta}$ and $v$ each have dimension $r$, we store a vector of dimension $p+2r$ for each image.
